### 2020-12-16 paper/sex_selection_spacing-ver5.tex
Add problem and questions to abstract

\noindent Over the past four decades, the Hindu women in India most likely to use 
sex-selective abortions---well-educated women with no sons---had the most substantial 
lengthening of birth intervals and the most biased sex ratios. 
As a result, we now see cases that reverse the traditional spacing pattern, with some
women with no sons having longer birth intervals than those with sons. 
Those least likely to use sex-selective abortions---less-educated women in rural 
areas---still follow the traditional pattern of short spacing when they have girls, with 
only limited evidence of sex selection. 
Because of the rapid lengthening in spacing, the standard fertility rates substantially 
overestimated how fast cohort fertility fell. 
Despite a convergence, cohort fertility is still 10\%--20\% higher than the fertility rate 
and above replacement level for all but the best-educated urban women. 
Infant mortality has declined substantially over time for all groups, with the fastest 
decline among the less educated. 
Short birth spacing is still associated with higher mortality, although considerably less 
so for the best-educated women. 
There is no evidence that repeated sex-selective abortions are associated with higher 
infant mortality for the child eventually born. 
Finally, it does not appear that the use of sex selection is declining.

% I show that, in India, the most substantial increases in Hindu women's
% birth interval over the last four decades came from the increasing use
% of sex-selective abortions. The women most likely to use sex
% selection---well-educated with no sons---had the most substantial
% lengthening of birth intervals and the most biased sex ratios.

@@

Second, what is the relationship between infant mortality and the changes in birth spacing 
and sex selection?
In India, birth intervals have traditionally been shorter with fewer sons, contributing 
to the higher mortality risk for girls
\citep{Whitworth2002,Bhalotra2008,Maitra2008,Jayachandran2011,Jayachandran2017a}.
Therefore, longer birth spacing, whether from sex selection or secular changes, may reduce 
mortality \citep{Conde-Agudelo2012,Molitoris2019}.
However, a counteracting effect is possible if longer birth intervals arise from multiple 
abortions because the short duration between pregnancies could increase mortality. 
% Nevertheless, there has been no attempt to examine how infant mortality has changed with 
% the increase in sex selection.

>>>>

% \noindent Over the past four decades, the Hindu women in India most likely to use 
% sex-selective abortions---well-educated women with no sons---had the most substantial 
% lengthening of birth intervals and the most biased sex ratios. 
% As a result, we now see cases that reverse the traditional spacing pattern, with some
% women with no sons having longer birth intervals than those with sons. 
% Those least likely to use sex-selective abortions---less-educated women in rural 
% areas---still follow the traditional pattern of short spacing when they have girls, with 
% only limited evidence of sex selection. 
% Because of the rapid lengthening in spacing, the standard fertility rates substantially 
% overestimated how fast cohort fertility fell. 
% Despite a convergence, cohort fertility is still 10\%--20\% higher than the fertility rate 
% and above replacement level for all but the best-educated urban women. 
% Infant mortality has declined substantially over time for all groups, with the fastest 
% decline among the less educated. 
% Short birth spacing is still associated with higher mortality, although considerably less 
% so for the best-educated women. 
% There is no evidence that repeated sex-selective abortions are associated with higher 
% infant mortality for the child eventually born. 
% Finally, it does not appear that the use of sex selection is declining.

\noindent India has experienced an increasingly male-biased sex ratio at birth following
the introduction of prenatal sex determination techniques. 
I examine three questions.
First, how did birth spacing respond to this and other secular developments? 
Second, did changes in birth spacing make us overestimate the decline in cohort fertility? 
Finally, did the advent of sex-selective abortions change the relationship between infant mortality and
birth spacing? 
Well-educated Hindu women with no sons had the most substantial lengthening of birth 
intervals and the most male-biased sex ratios, likely from sex-selective abortions. 
Hence, some women with no sons now have longer birth intervals than those with
sons, reversing the traditional spacing pattern. 
Less-educated women continued short spacing when they have girls, with only limited 
evidence of male-biased sex ratios. 
Because of the rapid lengthening in spacing, period fertility rates substantially 
overestimated how fast cohort fertility fell. 
Furthermore, predicted cohort fertility remains above replacement for all but the 
best-educated urban women. 
If lengthening of birth intervals occurs from repeated abortions, the associated short 
pregnancy spacing may counteract any positive effects of longer birth spacing. 
There is no evidence of this effect on infant mortality. 
Finally, judging from sex ratios, sex-selective abortion use is not declining.

@@

Second, what is the relationship between infant mortality and the changes in birth spacing
and sex selection? 
In India, birth intervals have traditionally been shorter with fewer sons, contributing to 
girls' higher mortality risk
\citep{Whitworth2002,Bhalotra2008,Maitra2008,Jayachandran2011,Jayachandran2017a}. 
Therefore, longer birth spacing may reduce mortality through, for example, diminished 
sibling competition \citep{Conde-Agudelo2012,Molitoris2019}. 
However, if the spacing between births lengthens because of sex-selective abortions, the 
spacing between pregnancies may still be very short. 
Short pregnancy spacing may lead to worse child outcomes because of maternal nutritional 
depletion and insufficient time to recover from the previous pregnancy. 
Hence, children born after long birth intervals where multiple abortions have punctuated 
those intervals may not see the same benefits as a child born after a long interval that 
is not punctuated by abortions.
